---
title: "farsFunctions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fars)
```


## Description

The farsFunctions functions extract state- and year-based data from input CSV file(s), creates corresponding file names,
verifies the "year" of the input file(s), summarizes the number of observations in each file and creates maps from the
input files.

  . fars_read() checks whether the CSV file exists. If the file exists, then it is read and coerced into a tibble.
  
  . make_filename() takes an input year, converts it to an integer, and then into a file name in the format: "accident_YYYY.csv.bz2",
    where YYYY represents the input year.
    
  . fars_read_years() checks whether the input vector "years", contains elements that appear in the output from make_filename() that
    produces file names in the format: "accident_YYYY.csv.bz2", where YYYY represents the year.
  
  . fars_summarize_years() computes the number of observations per month and year for each input file with format:"accident_YYYY.csv.bz2",
    where YYYY represents the year.
  
  . fars_map_state() creates a map of accidents that occurred within a specified state and during a specified year.



## Usage

fars_read("filename")

make_filename(year); make_filename("year"); make_filename(year)

fars_read_years(c(years))

fars_summarize_years(c(years))

fars_map_state(state.num, years)

#> Can access the file path like so:
file_path = system.file("extdata", "filename", package="farsFunctions")



## Arguments

filename        The name of the file that the function will read and coerce into a tibble.
state.num       The state code of interest.
year            The year of interest.
years           A vector of years.



## Value

  . fars_read() returns a tibble derived from the input CSV file.
  
  . make_filename() returns the file name in the format: "accident_YYYY.csv.bz2".
    
  . fars_read_years() returns a data frame with MONTH and year variables for each with a valid year.
  
  . fars_summarize_years() returns a tibble with the number of observations for each month and year for each valid input file.
  
  . fars_map_state() returns a map of accidents that occurred within a specified state and during a specified year.



## Examples

#### Reads the CSV file and converts to tibble
fars_read("accident_2013.csv.bz2")

#### Creates corresponding file name
make_filename(2013); make_filename("2014"); make_filename(2050)

#### Checks the validity of the year(s) of a vector of files
fars_read_years(c(2013, 2014, 2015))

#### Computes the number of observations per month and year for each input file
fars_summarize_years(c(2013, 2014, 2015))

#### Creates a map of accidents for specified state and year
fars_map_state(1, 2013)


-------------------------------------------------------------------------------------------------------------------------------
